LAGOS lake depth

### 2019-10-31 lagos_depth.Rmd
calculate depth classes

```{r num_lakes, eval = TRUE, echo=FALSE, message=FALSE}
dt_plot <- dt %>%
  st_drop_geometry() %>%
  dplyr::filter(!is.na(max_depth_m)) %>%
  group_by(state) %>%
  tally() %>%
  mutate(n_class = smart_cut(.$n, 5, "groups", 
                        labels = ~paste(sep = "-", round(.y[1]), round(.y[2])))) %>%
  right_join(usa_sf()) %>%
  st_set_geometry("geometry")

@@

```{r hypsography, eval=TRUE, fig.height = 3.2}
ids <- factor(c("convex", "concave"))
polys <- data.frame(
  id = rep(ids, each = 3),
  x = c(0, 100, 0, 
        0, 100, 100), 
  y = c(100, 0, 0, 
        100, 100, 0))

gg_hypso_ct <- ggplot(data = hypso_ct) +
  geom_polygon(data = polys, aes(x = x, y = y, group = id, fill = id), alpha = 0.2) +
  geom_line(aes(x = area_percent, y = depth_percent, group = llid)) +
  geom_abline(intercept = -100, slope = 1, color = "red") +
  ylim(c(100, 0)) +
  ylab("Depth (percent)") + xlab("Area (percent)") + 
  ggtitle(paste0("CT", " (", length(unique(hypso_ct$llid)), ") lakes")) + 
  theme(legend.position = c(0.84, 0.2), 
        legend.title = element_blank())

gg_hypso_mn <- ggplot(data = hypso_mn) +
  geom_polygon(data = polys, aes(x = x, y = y, group = id, fill = id), alpha = 0.2) +
  geom_line(aes(x = area_percent, y = depth_percent, group = llid)) +
  geom_abline(intercept = -100, slope = 1, color = "red") +
  ylim(c(100, 0)) +
  ylab("Depth (percent)") + xlab("Area (percent)") +
  ggtitle(paste0("MN", " (", length(unique(hypso_mn$llid)), ") lakes")) +
  theme(axis.title.y = element_blank(), 
        legend.position = "none")

plot_grid(gg_hypso_ct, gg_hypso_mn, 
rel_widths = c(1, 0.7))

@@

```{r classify, eval = TRUE}
# round the x value and test if y is greater or less than a straight-line curve
hypso_ct_classes <- hypso_ct %>%
  group_by(WBNAME) %>%
  summarize_all(median) %>%
  mutate(area_percent_rounded = round(area_percent)) %>%
  left_join(data.frame(y = rev(seq(0, 100, by = 1)), x = seq(0, 100, by = 1)), 
            by = c("area_percent_rounded" = "x")) %>%
  data.frame() %>%
  mutate(shape_class = case_when(depth_percent < y ~ "convex", 
                                 depth_percent > y ~ "concave", 
                                 TRUE ~ "neither")) %>%
  ungroup() %>%   dplyr::filter(!is.na(shape_class)) %>% 
    dplyr::select(WBNAME, llid, shape_class)

@@

```{r class-depths, eval = TRUE}

@@

>>>>

```{r num_lakes, eval=TRUE, echo=FALSE, message=FALSE}
dt_plot <- dt %>%
  st_drop_geometry() %>%
  dplyr::filter(!is.na(max_depth_m)) %>%
  group_by(state) %>%
  tally() %>%
  mutate(n_class = smart_cut(.$n, 5, "groups", 
                        labels = ~paste(sep = "-", round(.y[1]), round(.y[2])))) %>%
  right_join(usa_sf()) %>%
  st_set_geometry("geometry")

@@

```{r hypsography, eval=TRUE, fig.height = 2.8}
ids <- factor(c("convex", "concave"))
polys <- data.frame(
  id = rep(ids, each = 3),
  x = c(0, 100, 0, 
        0, 100, 100), 
  y = c(100, 0, 0, 
        100, 100, 0))

gg_hypso_ct <- ggplot(data = hypso_ct) +
  geom_polygon(data = polys, aes(x = x, y = y, group = id, fill = id), alpha = 0.2) +
  geom_line(aes(x = area_percent, y = depth_percent, group = llid)) +
  geom_abline(intercept = -100, slope = 1, color = "red") +
  ylim(c(100, 0)) +
  ylab("Depth (percent)") + xlab("Area (percent)") + 
  ggtitle(paste0("CT", " (n=", length(unique(hypso_ct$llid)), ") lakes [1]")) + 
  theme(legend.position = c(0.84, 0.2), 
        legend.title = element_blank())

gg_hypso_mn <- ggplot(data = hypso_mn) +
  geom_polygon(data = polys, aes(x = x, y = y, group = id, fill = id), alpha = 0.2) +
  geom_line(aes(x = area_percent, y = depth_percent, group = llid)) +
  geom_abline(intercept = -100, slope = 1, color = "red") +
  ylim(c(100, 0)) +
  ylab("Depth (percent)") + xlab("Area (percent)") +
  ggtitle(paste0("MN", " (n=", length(unique(hypso_mn$llid)), ") lakes [4]")) +
  theme(axis.title.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        axis.text.y = element_blank(),
        legend.position = c(0.72, 0.22), 
        legend.margin = margin(0, 4, 1, 1),
        legend.title = element_blank())

plot_grid(gg_hypso_ct + theme(legend.position = "none"), gg_hypso_mn, 
rel_widths = c(1, 0.88))

@@

```{r classify, eval=TRUE}
# round the x value and test if y is greater or less than a straight-line curve
hypso_ct_classes <- hypso_ct %>%
  group_by(WBNAME) %>%
  summarize_all(median) %>%
  mutate(area_percent_rounded = round(area_percent)) %>%
  left_join(data.frame(y = rev(seq(0, 100, by = 1)), x = seq(0, 100, by = 1)), 
            by = c("area_percent_rounded" = "x")) %>%
  data.frame() %>%
  mutate(shape_class = case_when(depth_percent < y ~ "convex", 
                                 depth_percent > y ~ "concave", 
                                 TRUE ~ "neither")) %>%
  ungroup() %>%   dplyr::filter(!is.na(shape_class)) %>% 
    dplyr::select(WBNAME, llid, shape_class)

@@

```{r class-depths, eval=TRUE}

@@

```{r depth-histogram, eval=TRUE}
# using Lauren's rounded (pretty) depth classes
dt2 <- read.csv("data/lagosus_depth.csv", stringsAsFactors = FALSE) %>%
  dplyr::filter(!is.na(max_depth_m)) %>%
  mutate(max_depth_m_class = smart_cut(max_depth_m, c(10, 30, 70, 200, Inf)))

table(dt2$max_depth_m_class)
round((table(dt2$max_depth_m_class) / nrow(dt2)) * 100, 2)
dplyr::filter(dt2, max_depth_m > 200)
```
